---
title: "Lab 03 - Nobel laureates"
author: "Jamieson Nathan"
date: "5/2/2025"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

```{r first_look}

dim(nobel)  # The first element is the number of rows and the second is the number of columns

head(nobel) # descriptions of rows

```
There appear to be 935 rows and 26 columns. As we can see in the above tibble, the rows refer to each individual Nobel laureate, and the columns give specific information like their name, year they won, what category it was, etc. 

### Exercise 2

```{r filtering_the_data}

library(dplyr)

nobel_living <- nobel %>%
  filter(is.na(died_date), !is.na(country), gender != "org")

nrow(nobel_living)

```
Confirmed that after filtering for currently living, country-available, and individual (not organisation) lareates, there are 228 remaining entries. 

### Exercise 3

```{r further_limits}

nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )

nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))

library(ggplot2)

ggplot(nobel_living_science, aes(x = country_us, fill = country_us)) +
  geom_bar() +
  facet_wrap(~ category, scales = "free_y") +
  coord_flip() +
  labs(title = "Nobel Prize Categories by Laureate's Country (USA vs Other)",
       x = "Country",
       y = "Count of Laureates",
       fill = "Country") +
  theme_minimal()

```

It appears that the buzzfeed article, remarkably, was accurate and that most Nobel Prize Laureates were in fact US-based when they won. 

### Exercise 4

```{r birth_country}

nobel_living <- nobel_living %>%
  mutate(born_country_us = if_else(born_country == "USA", "USA", "Other"))

nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics")) %>%
  mutate(born_country_us = if_else(born_country == "USA", "USA", "Other"))

nobel_living %>% 
  summarise(Count_US_Born = sum(born_country_us == "USA"))


```
By creating and summarizing this new variable, we find that 105 of the USA winners were born in the USA. 

### Exercise 5

```{r new_visual}

ggplot(nobel_living_science, aes(x = country_us, fill = born_country_us)) +
  geom_bar(position = "fill") +
  facet_wrap(~ category, scales = "free_y") +
  coord_flip() +
  scale_fill_manual(values = c("USA" = "turquoise", "Other" = "gold")) +
  labs(title = "Nobel Prize Categories by Laureate's Current and Birth Country",
       x = "Country of Prize Win",
       y = "Proportion of Laureates",
       fill = "Birth Country") +
  theme_minimal()

```
Based on the new data visual, it appears that it is accurate to say that some (perhaps not 'many') of the US-Based winners were immigrants or born elsewhere. 

### Exercise 6

```{r laureate_origins}

us_winners_foreign_born <- nobel_living %>%
  filter(country_us == "USA" & born_country_us == "Other") %>%
  count(born_country) %>%
  arrange(desc(n))

print(us_winners_foreign_born)

```
The most common appear to be the UK and Germany. 

In general, I actually struggle to see the why they use this data to support anti-immigration sentiment, as this pretty categorically shows that US-Immigrant laureates are from other Developed nations, where the petition and executive order this article is drumming support against is targeted at developing, Muslim nations. Nonetheless, an interesting exercise!

